Hydraulically-vulnerable trees survive on deep-water access in a tropical forest community

### 2020-10-26 manuscript.rmd
Fig. panel letters made lower case

# Materials and Methods
\
We developed a novel, empirical inverse model to estimate species effective rooting depths (ERD) for large trees of `r length(erd.sp)` species (Fig. \@ref(fig:schematic)), defined as the depth at which soil water potential best explained species-specific growth dynamics. The ERD model incorporated the impact of atmospheric drought on growth and was constrained with species-specific leaf vulnerability curves, by using existing data for eight species and developing trait-based proxies for the rest. To obtain the daily dynamics of soil water potential in the whole-soil column we locally parameterized and calibrated a 1-D hydrological water balance model. We evaluated ERD against existing deuterium to hydrogen content data for a subset of six tree species as an independent observation of ERD. We tested the relationship of ERD with above-ground hydraulic traits, namely, xylem vulnerability to embolism, branch hydraulic safety margins, leaf turgor loss point and maximum stem-specific hydraulic conductivity, for a subset of eight species with both datasets. Finally, we analysed species' mortality rates over 35 years to test whether ERD explained risk of drought-induced mortality.

@@

## Model ELM-FATES for soil matric potentials 
\
### Model description 
\
To estimate dynamics of water availability by depth over the forest’s rooting zone, we used the Energy Exascale Earth System Land Model (ELM), coupled with the Functionally Assembled Terrestrial Ecosystem Simulator (FATES) (hereafter, ELM-FATES). ELM is a land-model that, among many features, simulates the physics and conservative dynamics of water, energy and carbon fluxes. In particular, soil hydrologic fluxes are vertically resolved among discrete soil layers (1-D) similar to that in CLM4.5 [@Bonan:2013]. FATES is a community-based, open-source model used for studying climate-vegetation interactions. FATES is a vegetation demography model, with a size-structured group of plants (cohorts) and successional trajectory-based patches based on the ecosystem demography [@Moorcroft.2001] approach. FATES couples to ELM by a common interface of water and carbon fluxes. Detailed descriptions for ELM and FATES can be found elsewhere [@Fisher.2010; @Fisher.2015; @Koven:2020].

@@

## Model for effective rooting depth (ERD) 
\
### Model description 
\
To estimate effective rooting depth, we tested alternate empirical models for species 5-yearly census growth (Eq. \@ref(eq:M-chosen) and those in SI), which use effective rooting depth as an input. Many factors limit plant growth, including soil moisture, VPD, leaf area seasonality and radiation [@Lawrence.2019]. Because pre-dawn $\mathrm{\Psi}$~leaf~ approaches $\mathrm{\Psi}$~soil~, we used the latter in leaf hydraulic vulnerability curves to predict maximum diurnal $K$~leaf~, which in turn controls growth (see Table \@ref(tab:symbols) for trait symbols and their definitions). $K_\textrm{leaf}$ in our models thus incorporates the effects of rooting depth and rhizosphere water potential. This approach, however, does not account for the buffering effect of stem water storage capacitance on $\mathrm{\Psi}$~leaf~, and may thus overestimate ERD (see Discussion). VPD affects growth non-linearly, with growth increasing with VPD up-to a VPD threshold, then decreasing as leaf pores (stomata) close reducing water-uptake [@Grossiord.2020]. Apart from stomatal control, leaf-drop may further limit water-uptake and growth. We therefore regressed species-specific growth against multiplicative or additive effects of one or more factors calculated daily and averaged over 5-yearly censuses: VPD, $K_\textrm{leaf}$ and the leaf area index (LAI), all standardized between 0 and 1. Based on our preliminary analyses, additional incorporation of radiation did not improve model fitting. Due to the coincidence of higher temperature, higher radiation, and lower humidity during the dry season at BCI, the impact of VPD may as well be a combined impact of not only temperature and humidity, but also radiation.

@@

### Model parameterization 
\
#### Growth data

@@

\begin{equation}
\begin{split}
B_{s} = `r bcf[1]` `r bcf[2]`\textrm{WSG}_{s} + `r bcf[3]`\textrm{WSG}_{s}^2 
&`r bcf[4]`\textrm{WSG}_{s}\textrm{LMA}_{s} + `r bcf[5]`\textrm{LMA}_{s} `r bcf[6]`\textrm{LMA}_{s}^2 + \epsilon_{s,b}
(\#eq:b-mod)
\end{split}
\end{equation}
where $\epsilon_{s,a}$ and $\epsilon_{s,b}$ are error terms.

@@

# Results
## Soil water dynamics by depth
\
The 100 ELM-FATES ensemble members with the best fits to observed soil moisture dynamics captured soil moisture seasonality at multiple depths (Fig. \@ref(fig:hydro-calib)A, B), including the out of sample observations. These simulations also captured the dynamics in stream discharge (Fig. \@ref(fig:hydro-calib)C) and in evapotranspiration (Fig. \@ref(fig:hydro-calib)D), slightly underestimating peak discharge in wet years and slightly overestimating peak ET. The reduction in parameter range (Table S\@ref(tab:param))  in the best-fit ensembles compared to the tested global ranges showed that the model calibration was sensitive to the Ball-Berry stomatal slope parameter, the ELM root distribution parameter (that regulates the depth of the rooting profile), soil hydraulic conductivity profile, especially at depth, and the adjustment factor that varies soil hydraulic conductivity to account for macroporosity and direct flow paths (Fig. S\@ref(tab:sens-sm-et), S\@ref(tab:sens-run)).

@@

## Relationship of ERD with above-ground hydraulic traits
\
Species with deeper ERD showed greater maximum stem hydraulic conductance ($K_\textrm{max,stem}$, Spearman's' $r$ = `r cor.k`, $p$ = `r p.k`; Fig. \@ref(fig:erd-hyd)A), less negative leaf $\mathrm{\Psi}$~tlp~ ($r$ = `r cor.tlp`, $p$ = `r p.tlp`; Fig. \@ref(fig:erd-hyd)B), less negative $\mathrm{\Psi}$~88,stem~ and thus greater vulnerability to embolism from cavitation ($r$ = `r cor.stem.88`, $p$ = `r p.stem.88`; Fig. \@ref(fig:erd-hyd)C), and narrower hydraulic safety margins ($\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~; $r$ = `r cor.hsm`, $p$ = `r p.hsm`; Fig. \@ref(fig:erd-hyd)D).

@@

\
Figure \@ref(fig:hydro-calib): ELM-FATES calibration and evaluation. Observations (red lines) versus simulations (gray lines) from 100 best-fit ensemble-member runs of ELM-FATES are shown for daily Volumetric Water Content (VWC) by depth for horizontal time domain reflectometry (TDR) probes at three depths (0.1, 0.4 and 1 m; A) and a vertical probe (0-0.15 m; B); monthly stream discharge (Discharge, C) and monthly evapotranspiration from eddy flux tower (ET, D). Points (red) in (A) show average plot-wide observations of VWC with 95% CI (error bars). VWC data from the horizontal probes (lines in A) were not used for ELM-FATES calibration, but evaluation. Values in inset are average RMSE across the 100 best-fit simulations. For panel A, values at the top are for manual, plot-wide VWC and those at the bottom are for VWC from TDR probes. VWC is in units of cm^3^ cm^-3^, while ET and Discharge in mm month^-1^.

@@

\
Figure \@ref(fig:ab-fit): Parameters $A$ (panel A) & $B$ (panel B) for the leaf vulnerability curves that were fitted to observed data on $K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~ for `r length(sp.exp.param$sp)` species (Eq. \@ref(eq:vc-exp)) vs. those that were predicted from a set of models based on trait-proxies. (Eqs. \@ref(eq:a-mod)-\@ref(eq:b-mod)). Goodness-of-fit ($R$^2^) and $p$-values are given in inset.

@@

\
The $K_h$ segments were removed from the branches by repeatedly cutting back the branch on both ends while it was submerged in water to relax the hydraulic tension within the xylem ((REFERENCE)). Then the ends were shaved clean with a fresh razor. Segments were 11–34 cm in length (mean ± SD = 22 ± 4 cm) and 0.2–1.6 cm in diameter (0.7 ± 0.2 cm). The segments were placed in a hydraulic conductivity apparatus similar to that described by Sperry et al. (1998) except that flow rate was measured with graduated pipettes instead of a balance. The perfusion solution was 10 mM KCl filtered to 0.2 m. For each segment, we measured the flow rate at three pressure heads: two that ranged 0.5–2 kPa and at zero pressure. Following Torres-Ruiz et al. (2012), $K_h$ was calculated as the slope of the flow rate versus pressure gradient (i.e., pressure head divided by segment length) among the three pressure heads. Stem area specific hydraulic conductivity was ($K_\textrm{stem}$) was calculated by dividing $K_h$ by the cross-sectional area of stem. To calculate cross-sectional stem area, stem diameter was measured twice perpendicularly on both ends of the segment and averaged, then area was calculated assuming a circular cross section.

@@

```{r sens-sm-et, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "70%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-swp-all-depths}Parameter sensitivity of ELM-FATES soil water content (VWC) is shown for two sample months. (A, B)) as well as for ET for another two sample months (C, D), as sensitivity for a particular parameter varies depending on the chosen month. Parameters are color coded and are consistent across panels. Each point represents a simulation among the 5000-ensemble runs. See S\\@ref(tab:param) for description, global ranges and chosen ranges for parameters."}
knitr::include_graphics(file.path("figures/ELM_FATES_parameter_sensitivity/sens.sm.et.jpeg"))
```

```{r sens-run, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "70%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-swp-all-depths}Parameter sensitivity of ELM-FATES stream discharge (QRUNOFF) is shown for four sample months (A, B, C, D), as sensitivity for a particular parameter varies depending on the chosen month. Parameters are color coded and are consistent across panels. Each point represents a simulation among the 5000-ensemble runs. See S\\@ref(tab:param) for description, global ranges and chosen ranges for parameters."}
knitr::include_graphics(file.path("figures/ELM_FATES_parameter_sensitivity/sens.run.jpeg"))
```

```{r obs-plc-curves, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-obs-plc-curves}Vulnerability curves fitted to observed $K_\\textrm{leaf}$ vs. $\\mathrm{\\Psi}$~leaf~ data on ", length(sp.exp.param$sp)," tree species from BCI (A & B) with equation \\@ref(eq:vc-exp)--in absolute terms (A, B), as well as, in terms of loss of conductivity (%) (C, D). Each line represents a species, color coded by WSG (A, C), or LMA (B, D), with blue tones indicating smaller values of WSG or LMA and red tones indicating larger values.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/kmax_by_psi/Leaf/plot.obs.plc.jpeg"))
```

@@

#Supplementary Tables
\
```{r param, results = 'asis', echo = FALSE}
knitr::kable(param.table, caption = 'ELM-FATES parameters used to generate ensembles, with description, prescribed global ranges, rationale for the choice of ranges and references. Ranges for best-fit chosen ensembles are also given. Note that no. 1-6 are parameters specific to FATES.', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(param.table))))
```
#
```{r eddy, results = 'asis', echo = FALSE}
knitr::kable(eddy.table, caption = 'QA/QC procedure applied to eddy covariance fluxes. w: vertical wind speed; t: air temperature; q: water vapor density; c: CO~2~ density; u: horizontal wind speed. The bar indicates 30-min averaging and the carat indicates fluctuation ($\\overline{\\hat{x} = x - \\bar{x}}$)', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(eddy.table))))
```
#
```{r ab-param, results = 'asis', echo = FALSE}
knitr::kable(ab.table, caption = 'Leaf vulnerability curves parameters A & B, $K_\\textrm{max,leaf}$ and $\\mathrm{\\Psi}$ for the ", `r length(erd.sp)`, " species with ERD estimates. Species for which curves were fitted to observed data for $K_\\textrm{leaf}$ vs. $\\mathrm{\\Psi}$~leaf~ are identified with Source as Data, while species for which curves were obtained based on trait-proxies are identified with Source as Model', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(ab.table))))
```

>>>>

# Materials and Methods

We developed a novel, empirical inverse model to estimate species effective rooting depths (ERD) for large trees of `r length(erd.sp)` species (Fig. \@ref(fig:schematic)), defined as the depth at which soil water potential best explained species-specific growth dynamics. The ERD model incorporated the impact of atmospheric drought on growth and was constrained with species-specific leaf vulnerability curves, by using existing data for eight species and developing trait-based proxies for the rest. To obtain the daily dynamics of soil water potential in the whole-soil column we locally parameterized and calibrated a 1-D hydrological water balance model. We evaluated ERD against existing deuterium to hydrogen content data for a subset of six tree species as an independent observation of ERD. We tested the relationship of ERD with above-ground hydraulic traits, namely, xylem vulnerability to embolism, branch hydraulic safety margins, leaf turgor loss point and maximum stem-specific hydraulic conductivity, for a subset of eight species with both datasets. Finally, we analysed species' mortality rates over 35 years to test whether ERD explained risk of drought-induced mortality.

@@

## Model ELM-FATES for soil matric potentials 
### Model description 
\
To estimate dynamics of water availability by depth over the forest’s rooting zone, we used the Energy Exascale Earth System Land Model (ELM), coupled with the Functionally Assembled Terrestrial Ecosystem Simulator (FATES) (hereafter, ELM-FATES). ELM is a land-model that, among many features, simulates the physics and conservative dynamics of water, energy and carbon fluxes. In particular, soil hydrologic fluxes are vertically resolved among discrete soil layers (1-D) similar to that in CLM4.5 [@Bonan:2013]. FATES is a community-based, open-source model used for studying climate-vegetation interactions. FATES is a vegetation demography model, with a size-structured group of plants (cohorts) and successional trajectory-based patches based on the ecosystem demography [@Moorcroft.2001] approach. FATES couples to ELM by a common interface of water and carbon fluxes. Detailed descriptions for ELM and FATES can be found elsewhere [@Fisher.2010; @Fisher.2015; @Koven:2020].

@@

## Model for effective rooting depth (ERD)

### Model description 
\
To estimate effective rooting depth, we tested alternate empirical models for species 5-yearly census growth (Eq. \@ref(eq:M-chosen) and those in SI), which use effective rooting depth as an input. Many factors limit plant growth, including soil moisture, VPD, leaf area seasonality and radiation [@Lawrence.2019]. Because pre-dawn $\mathrm{\Psi}$~leaf~ approaches $\mathrm{\Psi}$~soil~, we used the latter in leaf hydraulic vulnerability curves to predict maximum diurnal $K$~leaf~, which in turn controls growth (see Table \@ref(tab:symbols) for trait symbols and their definitions). $K_\textrm{leaf}$ in our models thus incorporates the effects of rooting depth and rhizosphere water potential. This approach, however, does not account for the buffering effect of stem water storage capacitance on $\mathrm{\Psi}$~leaf~, and may thus overestimate ERD (see Discussion). VPD affects growth non-linearly, with growth increasing with VPD up-to a VPD threshold, then decreasing as leaf pores (stomata) close reducing water-uptake [@Grossiord.2020]. Apart from stomatal control, leaf-drop may further limit water-uptake and growth. We therefore regressed species-specific growth against multiplicative or additive effects of one or more factors calculated daily and averaged over 5-yearly censuses: VPD, $K_\textrm{leaf}$ and the leaf area index (LAI), all standardized between 0 and 1. Based on our preliminary analyses, additional incorporation of radiation did not improve model fitting. Due to the coincidence of higher temperature, higher radiation, and lower humidity during the dry season at BCI, the impact of VPD may as well be a combined impact of not only temperature and humidity, but also radiation.

@@

### Model parameterization

#### Growth data

@@

\begin{equation}
\begin{split}
B_{s} = `r bcf[1]` `r bcf[2]`\textrm{WSG}_{s} + `r bcf[3]`\textrm{WSG}_{s}^2 \\
&`r bcf[4]`\textrm{WSG}_{s}\textrm{LMA}_{s} + `r bcf[5]`\textrm{LMA}_{s} `r bcf[6]`\textrm{LMA}_{s}^2 + \epsilon_{s,b}
(\#eq:b-mod)
\end{split}
\end{equation}
where $\epsilon_{s,a}$ and $\epsilon_{s,b}$ are error terms.

@@

# Results
## Soil water dynamics by depth
\
The 100 ELM-FATES ensemble members with the best fits to observed soil moisture dynamics captured soil moisture seasonality at multiple depths (Fig. \@ref(fig:hydro-calib)a, b), including the out of sample observations. These simulations also captured the dynamics in stream discharge (Fig. \@ref(fig:hydro-calib)c) and in evapotranspiration (Fig. \@ref(fig:hydro-calib)d), slightly underestimating peak discharge in wet years and slightly overestimating peak ET. The reduction in parameter range (Table S\@ref(tab:param))  in the best-fit ensembles compared to the tested global ranges showed that the model calibration was sensitive to the Ball-Berry stomatal slope parameter, the ELM root distribution parameter (that regulates the depth of the rooting profile), soil hydraulic conductivity profile, especially at depth, and the adjustment factor that varies soil hydraulic conductivity to account for macroporosity and direct flow paths (Fig. S\@ref(fig:sens-sm-et), S\@ref(fig:sens-run)).

@@

## Relationship of ERD with above-ground hydraulic traits
\
Species with deeper ERD showed greater maximum stem hydraulic conductance ($K_\textrm{max,stem}$, Spearman's' $r$ = `r cor.k`, $p$ = `r p.k`; Fig. \@ref(fig:erd-hyd)a), less negative leaf $\mathrm{\Psi}$~tlp~ ($r$ = `r cor.tlp`, $p$ = `r p.tlp`; Fig. \@ref(fig:erd-hyd)b), less negative $\mathrm{\Psi}$~88,stem~ and thus greater vulnerability to embolism from cavitation ($r$ = `r cor.stem.88`, $p$ = `r p.stem.88`; Fig. \@ref(fig:erd-hyd)c), and narrower hydraulic safety margins ($\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~; $r$ = `r cor.hsm`, $p$ = `r p.hsm`; Fig. \@ref(fig:erd-hyd)d).

@@

\
Figure \@ref(fig:hydro-calib): ELM-FATES calibration and evaluation. Observations (red lines) versus simulations (gray lines) from 100 best-fit ensemble-member runs of ELM-FATES are shown for daily Volumetric Water Content (VWC) by depth for horizontal time domain reflectometry (TDR) probes at three depths (0.1, 0.4 and 1 m; a) and a vertical probe (0-0.15 m; b); monthly stream discharge (Discharge, c) and monthly evapotranspiration from eddy flux tower (ET, d). Points (red) in (a) show average plot-wide observations of VWC with 95% CI (error bars). VWC data from the horizontal probes (lines in a) were not used for ELM-FATES calibration, but evaluation. Values in inset are average RMSE across the 100 best-fit simulations. For panel a, values at the top are for manual, plot-wide VWC and those at the bottom are for VWC from TDR probes. VWC is in units of cm^3^ cm^-3^, while ET and Discharge in mm month^-1^.

@@

\
Figure \@ref(fig:ab-fit): Parameters $A$ (panel a) & $B$ (panel b) for the leaf vulnerability curves that were fitted to observed data on $K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~ for `r length(sp.exp.param$sp)` species (Eq. \@ref(eq:vc-exp)) vs. those that were predicted from a set of models based on trait-proxies. (Eqs. \@ref(eq:a-mod)-\@ref(eq:b-mod)). Goodness-of-fit ($R$^2^) and $p$-values are given in inset.

@@

\
The $K_h$ segments were removed from the branches by repeatedly cutting back the branch on both ends while it was submerged in water to relax the hydraulic tension within the xylem ((REFERENCE)). Then the ends were shaved clean with a fresh razor. Segments were 11–34 cm in length (mean ± SD = 22 ± 4 cm) and 0.2–1.6 cm in diameter (0.7 ± 0.2 cm). The segments were placed in a hydraulic conductivity apparatus similar to that described by Sperry et al. (1998) except that flow rate was measured with graduated pipettes instead of a balance. The perfusion solution was 10 mM KCl filtered to 0.2 mM. For each segment, we measured the flow rate at three pressure heads: two that ranged 0.5–2 kPa and at zero pressure. Following Torres-Ruiz et al. (2012), $K_h$ was calculated as the slope of the flow rate versus pressure gradient (i.e., pressure head divided by segment length) among the three pressure heads. Stem area specific hydraulic conductivity was ($K_\textrm{stem}$) was calculated by dividing $K_h$ by the cross-sectional area of stem. To calculate cross-sectional stem area, stem diameter was measured twice perpendicularly on both ends of the segment and averaged, then area was calculated assuming a circular cross section.

@@

```{r sens-sm-et, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "70%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-sens-sm-et}Parameter sensitivity of ELM-FATES soil water content (VWC) is shown for two sample months (a, b) as well as for ET for two other sample months (c, d), as sensitivity for a particular parameter varies depending on the chosen month. Parameters are color coded and are consistent across panels. Each point represents a simulation among the 5000-ensemble runs. See S\\@ref(tab:param) for description, global ranges and chosen ranges for parameters."}
knitr::include_graphics(file.path("figures/ELM_FATES_parameter_sensitivity/sens.sm.et.jpeg"))
```

```{r sens-run, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "70%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-sens-run}Parameter sensitivity of ELM-FATES stream discharge (QRUNOFF) is shown for four sample months (a-d), as sensitivity for a particular parameter varies depending on the chosen month. Parameters are color coded and are consistent across panels. Each point represents a simulation among the 5000-ensemble runs. See S\\@ref(tab:param) for description, global ranges and chosen ranges for parameters."}
knitr::include_graphics(file.path("figures/ELM_FATES_parameter_sensitivity/sens.run.jpeg"))
```

```{r obs-plc-curves, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-obs-plc-curves}Vulnerability curves fitted to observed $K_\\textrm{leaf}$ vs. $\\mathrm{\\Psi}$~leaf~ data on ", length(sp.exp.param$sp)," tree species from BCI (a, b) with equation \\@ref(eq:vc-exp)--in absolute terms (a, b), as well as, in terms of loss of conductivity (%) (c, d). Each line represents a species, color coded by WSG (a, c), or LMA (b, d), with blue tones indicating smaller values of WSG or LMA and red tones indicating larger values.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/kmax_by_psi/Leaf/plot.obs.plc.jpeg"))
```

@@

#Supplementary Tables
\
```{r param, results = 'asis', echo = FALSE}
knitr::kable(param.table, caption = 'ELM-FATES parameters used to generate ensembles, with description, prescribed global ranges, rationale for the choice of ranges and references. Ranges for best-fit chosen ensembles are also given. Note that no. 1-6 are parameters specific to FATES.', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(param.table))))
```
#
```{r eddy, results = 'asis', echo = FALSE}
knitr::kable(eddy.table, caption = 'QA/QC procedure applied to eddy covariance fluxes. w: vertical wind speed; t: air temperature; q: water vapor density; c: CO~2~ density; u: horizontal wind speed. The bar indicates 30-min averaging and the carat indicates fluctuation ($\\overline{\\hat{x} = x - \\bar{x}}$)', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(eddy.table))))
```
#
```{r ab-param, results = 'asis', echo = FALSE}
knitr::kable(ab.table, caption = paste0("\\label{lab-ab-param}Leaf vulnerability curves parameters *A* & *B*, $K_\\textrm{max,leaf}$ and $\\mathrm{\\Psi}$ for the ", length(erd.sp), " species with ERD estimates. Species for which curves were fitted to observed data for $K_\\textrm{leaf}$ vs. $\\mathrm{\\Psi}$~leaf~ are identified with Source as Data, while species for which curves were obtained based on trait-proxies are identified with Source as Model."), booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(ab.table))))
```
